Parallel Information Set Generation for Kriegspiel

### 2012-09-15 sbac/paper.tex
Several minor typos and mechanics issues addressed.

\section{Introduction}
In imperfect information games, players do not have access to full knowledge of
the world. Examples of imperfect information include hidden cards in
Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden
tiles in Scrabble~\cite{richards07opponent}, or hidden pieces in 
Kriegspiel (partially observable chess)~\cite{li94chess}. The game tree nodes
which are indistinguishable to a player since they differ only in the
information that is hidden to the player by rule compose that player's {\em
information set.}  The ability to estimate the value of the possible states and
to reason about the probability distribution over those states is crucial to
playing imperfect information games well~\cite{schofield12hyperplay}.

@@

The listing of a sequence of moves from the perspective of one player lists all of that player's moves exactly and
includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
players}, that are consistent with those observations.  In other words, any sequence of moves that would have generated
the same sequence of observations for Black is in Black's information set.  Figure~\ref{abbrevoutput} shows a portion of the
information set for Black after both player's have made five moves.  From this list of possibilities, Black can infer
that White must have a bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with
Black's observations up to that point.

@@

\textbf{The Information Set Generation}
(ISG) operation can be formulated as a combinatorial search problem.  The input to the algorithm is a list
of $t$ observations for one player $k$ (denoted $o_{1:t}^{(k)}$ in a Kriegspiel game.    
\begin{figure}[htpb]
\small
%\begin{algorithm}
\begin{boxedminipage}{\columnwidth}
\begin{algorithmic}[1]
\Function{DFS-Infoset}{$a_{1:t}$, $o_{1:T}^{(k)}$, $I$, $s$, $t$}
  \If {$t > T$}
    \State {$I \leftarrow I \cup a_{1:t}$} %\Comment{$a_{1:t}$ leads to $o_{1:T}$ for $P_k$; add it to the information set.}
  \Else {
    \For{\textbf{each} legal action $a$ at $s$}
      \State {$s'$ = {\sc GetNextState}($s$,$a$)}\label{line:update}
      \If {$s'.o_{t+1}^{(k)}=o_{t+1}^{(k)}$}
        \State {{\sc DFS-Infoset}($a_{1:t}a$, $o_{1:T}^{(k)}$, $I$, $s'$, $t+1$)} 
          %\Comment{ Recall that $N(a_{1:t}a) = T(T(a_{1:t}),a)$}
      \EndIf
    \EndFor
  }\EndIf
\EndFunction
\end{algorithmic}
\end{boxedminipage}
\caption{Core of Information Set Generation Algorithm}
\label{codelisting1}
\vspace{-0.2in}
\end{figure}

@@

\section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
We built our parallel ISG search engine on top of the {\sc Charm++} run-time
system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
is an extension to the C++ programming language that provides
machine-independent infrastructure for parallel computation.  The language and
its accompanying runtime system have been ported to many shared-memory and
distributed-memory platforms. {\sc Charm++} is used by many highly scalable scientific applications such as NAMD~\cite{NAMDIPDPS08} and ChaNGa~\cite{2007_ChaNGaScaling}

@@

Another strategy we evaluated is the neighbor based load balancing which imposes a virtual topology on the processors and load exchange occur between neighbors only. In this strategy, the load balancing is initiated by overloaded processors. When a processor becomes overloaded, it sends work to its neighbors.

@@

We next study the variation in the efficiency of the parallelization of the
algorithm as size of the problem increases.  Here, we varied the number
of plies in the test instance from 8 to 12. We used the work stealing
load-balancing strategy.  
\begin{figure}[ht] 
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthspe.pdf}
\end{minipage}
\caption{(Top) Execution time for three different sizes of problem as the number of cores used increases.
(Bottom) Parallel Efficiency (wrt to 16 cores) vs number of cores, to help see the effect of problem size and iso-efficiency.}
\label{3depths}
\vspace{-0.3in}
\end{figure}
The results are shown in Figure~\ref{3depths}.  In the easiest case (depth 8),
there is not enough additional parallel work to justify the jump from 256 to
1,024 processors, and performance actually degrades.  Aside from this anomaly,
however, we see consistent speedups as we increase the number of processors. 
The plot on the bottom helps to illustrate the effect of problem size on parallel efficiency (defined as Speedup/P) and helps understand the iso-efficiency of the problem. Iso-efficiency~\cite{b:grama-isoefficiency} means how should the problem size be increased with increasing number of processors to maintain same level of parallel efficiency. As the size of the problem increases, the amount of parallelizable work available also increases, and hence parallel efficiency is higher for larger problem sizes and constant number of processors (Figure~\ref{3depths} bottom). As the number of processors increases, parallel efficiency decreases for constant problem size. Figure~\ref{3depths} suggests that in parallel information set generation, to maintain constant parallel efficiency, problem size (search depth) should be roughly increased by two when the number of processor in increased by 4 times.

@@

\begin{figure}[t]
 \centering \includegraphics[scale=1]{plots/new.pdf} %
\caption{Performance results for Information Set Generation for two games of Kriegspiel, position ten}
\label{prob4}
\vspace{-0.2in} 
\end{figure}

>>>>

\section{Introduction}
In imperfect information games, players do not have access to full knowledge of
the world. Examples of imperfect information include hidden cards in
Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden
tiles in Scrabble~\cite{richards07opponent}, or hidden pieces in 
Kriegspiel (partially observable chess)~\cite{li94chess}. The game tree nodes
that are indistinguishable to a player since they differ only in the
information that is hidden to the player by rule compose that player's {\em
information set.}  The ability to estimate the value of the possible states and
to reason about the probability distribution over those states is crucial to
playing imperfect information games well~\cite{schofield12hyperplay}.

@@

The listing of a sequence of moves from the perspective of one player shows all of that player's moves exactly and
includes the declarations from the referee.  We refer to such a list as a player's observations.  Given a list of
observations for a sequence of moves, an information set is the set of all possible {\em sequences of moves for both
players}, that are consistent with those observations.  In other words, any sequence of moves that would have generated
the same sequence of observations for Black is in Black's information set.  Figure~\ref{abbrevoutput} shows a portion of the
information set for Black after both players have made five moves.  From this list of possibilities, Black can infer
that White must have a bishop at g5 and a pawn at d6.  There are 25 possible sequences of moves that are consistent with
Black's observations up to that point.

@@

\textbf{The Information Set Generation}
(ISG) operation can be formulated as a combinatorial search problem.  The input to the algorithm is a list
of $t$ observations for one player $k$ (denoted $o_{1:t}^{(k)}$) in a Kriegspiel game.    
\begin{figure}[htpb]
\small
%\begin{algorithm}
\begin{boxedminipage}{\columnwidth}
\begin{algorithmic}[1]
\Function{DFS-Infoset}{$a_{1:t}$, $o_{1:T}^{(k)}$, $I$, $s$, $t$}
  \If {$t > T$}
    \State {$I \leftarrow I \cup a_{1:t}$} %\Comment{$a_{1:t}$ leads to $o_{1:T}$ for $P_k$; add it to the information set.}
  \Else {
    \For{\textbf{each} legal action $a$ at $s$}
      \State {$s'$ = {\sc GetNextState}($s$,$a$)}\label{line:update}
      \If {$s'.o_{t+1}^{(k)}=o_{t+1}^{(k)}$}
        \State {{\sc DFS-Infoset}($a_{1:t}a$, $o_{1:T}^{(k)}$, $I$, $s'$, $t+1$)} 
          %\Comment{ Recall that $N(a_{1:t}a) = T(T(a_{1:t}),a)$}
      \EndIf
    \EndFor
  }\EndIf
\EndFunction
\end{algorithmic}
\end{boxedminipage}
\caption{Core of Information Set Generation Algorithm}
\label{codelisting1}
\vspace{-0.2in}
\end{figure}

@@

\section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
We built our parallel ISG search engine on top of the {\sc Charm++} run-time
system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
is an extension to the C++ programming language that provides
machine-independent infrastructure for parallel computation.  The language and
its accompanying runtime system have been ported to many shared-memory and
distributed-memory platforms. {\sc Charm++} is used by many highly scalable scientific applications such as NAMD~\cite{NAMDIPDPS08} and ChaNGa~\cite{2007_ChaNGaScaling}.

@@

Another strategy we evaluated is the neighbor-based load balancing which imposes a virtual topology on the processors and load exchange occur between neighbors only. In this strategy, the load balancing is initiated by overloaded processors. When a processor becomes overloaded, it sends work to its neighbors.

@@

We next study the variation in the efficiency of the parallelization of the
algorithm as size of the problem increases.  Here, we varied the number
of plies in the test instance from 8 to 12. We used the work stealing
load-balancing strategy.  
\begin{figure}[ht] 
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthspe.pdf}
\end{minipage}
\caption{(Top) Execution time for three different sizes of problem as the number of cores used increases.
(Bottom) Parallel Efficiency (wrt to 16 cores) vs number of cores, to help see the effect of problem size and iso-efficiency.}
\label{3depths}
\vspace{-0.3in}
\end{figure}
The results are shown in Figure~\ref{3depths}.  In the easiest case (depth 8),
there is not enough additional parallel work to justify the jump from 256 to
1,024 processors, and performance actually degrades.  Aside from this anomaly,
however, we see consistent speedups as we increase the number of processors. 
The plot on the bottom helps to illustrate the effect of problem size on parallel efficiency (defined as Speedup/P) and helps understand the iso-efficiency of the problem. Iso-efficiency~\cite{b:grama-isoefficiency} quantifies how the problem size should be increased with increasing number of processors to maintain same level of parallel efficiency. As the size of the problem increases, the amount of parallelizable work available also increases, and hence parallel efficiency is higher for larger problem sizes and constant number of processors (Figure~\ref{3depths} bottom). As the number of processors increases, parallel efficiency decreases for constant problem size. Figure~\ref{3depths} suggests that in parallel information set generation, to maintain constant parallel efficiency, problem size (search depth) should be roughly increased by two when the number of processors is increased by 4 times.

@@

\begin{figure}[t]
 \centering \includegraphics[scale=1]{plots/new.pdf} %
\caption{Performance results for Information Set Generation for two games of Kriegspiel, position ten.}
\label{prob4}
\vspace{-0.2in} 
\end{figure}
